### 2022-08-11 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@3311c55a5d9572bc262b0c1d5d170f7d02afe1ed 🚀

Our models need information on the number of species and the number of
interactions. However, since the later is rarely estimated empirically,
ecologists might need to use predictive methods to estimate the total
number of interactions in a food web.

@@

The flexible links model is a generative model, i.e.~it can generate
plausible values of the predicted variable. We thus simulated \(1000\)
values of \(L\) for different values of \(S\) using the joint posterior
distribution of our model parameters (eq.~\ref{eq:BBpost}), and
calculated the mean degree for each simulated values. The resulting
distributions are shown in the left panel of
fig.~\ref{fig:degree_dist_fl} for three different values of species
richness. In the right panel of fig.~\ref{fig:degree_dist_fl}, we show
how the probability distribution for the mean degree constraints can be
used to generate a distribution of maximum entropy degree distributions,
since each simulated value of mean degree generates a different maximum
entropy degree distribution (eq.~\ref{eq:lagrange_dd} and
eq.~\ref{eq:lagrange2_dd}).

\begin{figure}
\hypertarget{fig:degree_dist_fl}{%
\centering
\includegraphics{figures/maxent_degree_dist_fl.png}
\caption{(a) Probability density of the predicted mean degree for
different values of species richness. The number of interactions was
predicted using the flexible links model fitted to all empirical
networks in our complete dataset (eq.~\ref{eq:BBpost}). (b) Degree
distributions of maximum entropy for a network of \(S=27\) species and
different numbers of interactions \(L\). The numbers of interactions
correspond to the lower and upper bounds of the \(67%
\), \(89%
\), and \(97%
\) percentile intervals (PI), as well as the median (in blue), of the
counterfactuals of the flexible links model. Each degree distribution
was obtained using eq.~\ref{eq:lagrange_dd} after solving numerically
eq.~\ref{eq:lagrange2_dd} using different values for the mean degree
constraint \(2L/S\).}\label{fig:degree_dist_fl}
}
\end{figure}

@@

Our heuristic MaxEnt models, predicting adjacency matrices, were
compared with two topological null models. The first is the type I null
model of Fortuna and Bascompte (2006), in which the probability that a
species \(i\) predates on another species \(j\) is given by

@@

where \(k_{in}\) and \(k_{out}\) are the in and out-degrees,
respectively. The type I null model is based on connectance, whereas the
type II null model is based on the joint degree sequence. Therefore, the
type I and II topological null models correspond with our type I and II
MaxEnt network models, respectively, since they use similar constraints.

We predicted both types of null networks for all empirical networks in
our complete dataset (\(n = 257\)). We converted all probabilistic
networks to Boolean networks by generating \(100\) random Boolean
networks for each of these probabilistic webs. Then, we counted the
number of times each interaction was sampled, and kept the \(L\) entries
that were drawn the most amount of time, with \(L\) given by the number
of interactions in each food web. This ensured that the resulting null
networks had the same number of interactions as their empirical
counterparts.

@@

We also compared our MaxEnt network models with a neutral model of
relative abundances, in which the probabilities of interaction are given
by

@@

where \(n_i\) and \(n_j\) are the abundances (or biomass) of both
species, and \(N\) is the total abundance (or biomass) of all species in
the network. We predicted neutral abundance matrices for all empirical
networks in our abundance dataset (\(n = 19\)), and converted these
weighted matrices to Boolean networks using an approach analogue to the
one we used for our null models.

>>>>

Our analytical MaxEnt models require information on the number of
species and the number of interactions. However, since the later is
rarely measured empirically, ecologists might need to use predictive
models to estimate the total number of interactions in a food web before
using MaxEnt. Here we illustrate how this can be done by combining both
models sequentially.

@@

The flexible links model is a generative model, i.e.~it can generate
plausible values of the predicted variable. We thus simulated \(1000\)
values of \(L\) for different values of \(S\) using the joint posterior
distribution of our model parameters (eq.~\ref{eq:BBpost}), and
calculated the mean degree for each simulated value. The resulting
distributions are shown in the left panel of
fig.~\ref{fig:degree_dist_fl} for three different values of species
richness. In the right panel of fig.~\ref{fig:degree_dist_fl}, we show
how the probability distribution for the mean degree constraints can be
used to generate a distribution of maximum entropy degree distributions,
since each simulated value of mean degree generates a different maximum
entropy degree distribution (eq.~\ref{eq:lagrange_dd} and
eq.~\ref{eq:lagrange2_dd}).

\begin{figure}
\hypertarget{fig:degree_dist_fl}{%
\centering
\includegraphics{figures/maxent_degree_dist_fl.png}
\caption{(a) Probability density of the mean degree of a food web
obtained using different values of species richness \(S\). The number of
interactions \(L\) was simulated \(1000\) times using the flexible links
model fitted to all empirical networks in our complete dataset
(eq.~\ref{eq:BBpost}). The mean degrees \(2L/S\) were then obtained from
these simulated values. (b) Degree distributions of maximum entropy for
a network of \(S=27\) species and different numbers of interactions. The
numbers of interactions correspond to the lower and upper bounds of the
\(67%
\), \(89%
\), and \(97%
\) percentile intervals (PI), as well as the median (in blue), of the
counterfactuals of the flexible links model. Each degree distribution of
maximum entropy was obtained using eq.~\ref{eq:lagrange_dd} after
solving numerically eq.~\ref{eq:lagrange2_dd} using different values of
the mean degree constraint.}\label{fig:degree_dist_fl}
}
\end{figure}

@@

The predictions of our heuristic maximum entropy models were compared
against two topological null models. These null models use the same
ecological information as our heuristic models and thus constitute an
adequate baseline for comparison. The first is the type I null model of
Fortuna and Bascompte (2006), in which the probability that a species
\(i\) predates on another species \(j\) is given by

@@

where \(k_{in}(j)\) and \(k_{out}(i)\) are the in and out-degrees of
species \(j\) and \(i\), respectively. The type I null model is based on
connectance, whereas the type II null model is based on the joint degree
sequence. Therefore, the type I and II topological null models
correspond to our type I and II heuristic MaxEnt models, respectively,
since they use similar constraints.

We generated probabilistic networks using both types of null models for
all empirical food webs in our complete dataset. Then, we converted
these networks to adjacency matrices of Boolean values by generating
\(100\) random networks for each of these probabilistic webs. Finally,
we kept the \(L\) entries that were sampled the most amount of times,
with \(L\) given by the number of interactions in each food web. This
ensured that the resulting null networks had the same number of
interactions as their empirical counterparts. Thus, for each null model,
we ended up with one null adjacency matrix for each empirical network.

@@

We also compared our heuristic MaxEnt models with a neutral model of
relative abundances, in which the probabilities of interaction are given
by

@@

where \(n_i\) and \(n_j\) are the abundances (or biomass) of both
species, and \(N\) is the total abundance (or biomass) of all species in
the network. We predicted neutral abundance matrices for all empirical
food webs in our abundance dataset, and converted these weighted
networks to adjacency matrices of Boolean values using the same method
as the one we used for our null models.

### 2022-08-11 suppmat.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@3311c55a5d9572bc262b0c1d5d170f7d02afe1ed 🚀

\begin{figure}
\hypertarget{fig:heatmap}{%
\centering
\includegraphics{figures/heatmap_disconnected.png}
\caption{Probability that a species is isolated in its food web
according to the degree distribution of maximum entropy. We derived
degree distributions of maximum entropy given a range of values of \(S\)
and \(L\), and plotted the probability that a species has a degree \(k\)
of \(0\) (log-scale color bar). Here species richness varies between
\(5\) and \(100\) species, by increment of \(5\) species. For each level
of species richness, the numbers of links correspond to all 20-quantiles
of the interval between \(0\) and \(S^2\). The black line marks the
\(S-1\) minimum numbers of links required to have no isolated
species.}\label{fig:heatmap}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:entropy_size}{%
\centering
\includegraphics{figures/difference_entropy.png}
\caption{Difference in SVD entropy between MaxEnt (type II MaxEnt
network model) and empirical food webs as a function of (a) species
richness, (b) the number of links, and (c) connectance. Empirical
networks include most food webs archived on Mangal, as well as the New
Zealand and Tuesday lake datasets (our complete dataset). Maximum
entropy networks were derived using a simulating annealing algorithm to
find the network of maximum SVD entropy while maintaining the joint
degree sequence. Regression lines are plotted in each
panel.}\label{fig:entropy_size}
}
\end{figure}

>>>>

\begin{figure}
\hypertarget{fig:heatmap}{%
\centering
\includegraphics{figures/heatmap_disconnected.png}
\caption{Probability that a species is isolated in its food web
according to the degree distribution of maximum entropy. We derived many
degree distributions of maximum entropy given a range of values of \(S\)
and \(L\), and plotted the probability that a species has a degree \(k\)
of \(0\) (log-scale color bar). Here species richness varies between
\(5\) and \(100\) species, by increment of \(5\) species. For each level
of species richness, the numbers of interactions correspond to all
20-quantiles of the interval between \(0\) and \(S^2\). The black line
marks the \(S-1\) minimum numbers of interactions required to have no
isolated species.}\label{fig:heatmap}
}
\end{figure}

@@

\begin{figure}
\hypertarget{fig:entropy_size}{%
\centering
\includegraphics{figures/difference_entropy.png}
\caption{Difference in SVD entropy between MaxEnt (type II MaxEnt
network model) and empirical food webs as a function of (a) species
richness, (b) the number of interactions, and (c) connectance. Empirical
networks include most food webs archived on Mangal, as well as the New
Zealand and Tuesday lake datasets (our complete dataset). Maximum
entropy networks were derived using a simulating annealing algorithm to
find the network of maximum SVD entropy while maintaining the joint
degree sequence. Regression lines are plotted in each
panel.}\label{fig:entropy_size}
}
\end{figure}
